---
title: "Handling FARS Data"
author: "Jens Ulrik Hansen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Handling FARS Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

---

This package is designed to read in FARS data, that is, *Fatality Analysis Reporting System* data from the US National Highway Traffic Safety Administration, and do simple summaries and plots of this data.

The package contains two main functions:

- fars_summarize_years,
- fars_map_state,

in addition to the three helper functions "fars_read", "make_filename", and "fars_read_years". The function "make_filename" takes a year and returns the name of the file containing the FARS data for that year. The function "fars_read" simply reads a csv file from a file name (as the one created from the "make_filename" function just mentioned) and puts the data into a tbl_df data frame. Finally, the function "fars_read_years",  takes a vector of years and return a list of data frames with month and years for the FARS data for the input vector of years.

In the following sections the two main functions are described in more details in the order of their intented use.

## fars_summarize_years

This function take a vector of years and returns a summary of how many data point that are for each year and month in the given range of years. For instance, the input
```{r, eval=FALSE}
fars_summarize_years(2013:2015)
```
returns:
```{r, echo=FALSE}
structure(list(MONTH = 1:12, `2013` = c(2230L, 1952L, 2356L, 
2300L, 2532L, 2692L, 2660L, 2899L, 2741L, 2768L, 2615L, 2457L
), `2014` = c(2168L, 1893L, 2245L, 2308L, 2596L, 2583L, 2696L, 
2800L, 2618L, 2831L, 2714L, 2604L), `2015` = c(2368L, 1968L, 
2385L, 2430L, 2847L, 2765L, 2998L, 3016L, 2865L, 3019L, 2724L, 
2781L)), .Names = c("MONTH", "2013", "2014", "2015"), class = c("tbl_df", 
"tbl", "data.frame"), row.names = c(NA, 12L))
```


## fars_map_state

The "fars_map_state" function takes a state (as a numeric value) and a yea and plots the accident that happens that year in that state. For istance,
```{r, eval=FALSE}
fars_map_state(6, 2013)
```
plots all the accidents in California for the year 2013 on a simple Calofornia state map.
